Radial and rotational velocities of young brown dwarfs and very low-mass stars in the Upper Scorpius OB association and the $\rho$~Ophiuchi cloud core

### 2006-06-12 BD_radvel/tmp.tex
Minor changes.
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@739 25464f21-6099-402c-b30a-640f1d40d38b

The radial velocities of each object were determined by using the cross-correlation
function of the object spectrum with that of a template star which has
a similar spectra type. By visual inspection,
the wavelength ranges used for the cross-correlation calculations
are chosen by avoiding the regions of spectra affected by the telluric
lines, and defects and fringes (in near infrared) of the CCDs. The
radial velocities of objects with respect to the template are obtained
by measuring the location of the peak in the cross-correlation
function. The location of the peak is determined by fitting the
cross-correlation function by a function which consists of two gaussians 
(with a common centre) plus a constant term. LHS~49 (Proxima Cen, M5.5) was
chosen as the template for this purpose. The radial velocity of the
template object LHS~49 was obtained by measuring the wavelength shifts
of the prominent photospheric absorption features \ion{K}{i}~$\lambda\lambda$7664.911,~7698.974.
This gives us $\mathrm{RV_{LHS49}=}-22.6\pm0.5\,\kmps$, which is
in good agreement with the earlier measurement of \citet{garcia-sanchez:2001}
who found $\mathrm{RV_{LHS49}=}-21.7\pm1.8\,\kmps$. The heliocentric
RV of each object can be then calculated by adding $\mathrm{RV_{LHS49}}$
with the RV of each object with respect to LHS~49. In the following
measurements of the heliocentric radial velocities, our measurement
($\mathrm{RV_{LHS49}=}-22.6\pm0.5\,\kmps$) will be used for consistency.

Before applying the cross-correlation technique to our main targets,
we have applied the technique to the radial velocity standard HD~140538
for which an high-accuracy RV measurement via the fixed-configuration,
cross-dispersed échelle spectrograph Elodie \citep{baranne:1996}
is available. This was done so to ensure not only the validity of
the cross-correlation technique, but also the validity of the wavelength
calibration. In this test, we found the heliocentric $\mathrm{RV_{HD\,140538}=}18.8\pm0.6\,\kmps$
which is in good agreement with the Elodie radial velocity measurement
of $19.00\pm0.05\,\mathrm{km\, s^{-1}}$ \citep{udry:1999}.

The result of the heliocentric RV measurements (from two epochs for
each object) is summarised in Table~\ref{tab:summary_results} along
with the uncertainties. The table also lists the uncertainties in
the `relative' radial velocities ($\sigma_{\mathrm{RRV}}$) which
is the uncertainty of RV with respect to the template star. The
two consecutive measurements of RVs (from same nights) are used to
find the (absolute/heliocentric) RV of the night, and their uncertainties are obtained from
the standard deviations of the mean ($\sigma_{\mathrm{RRV}}$) and
the uncertainty of the template RV value (c.f.~\citealt{joergens:2006}).
Note that for finding a shift in the radial velocities from two
different epochs, only $\sigma_{\mathrm{RRV}}$ is important since one
only requires relative (with respect to a template) RV values.
The `average' radial velocities $\left (\overline{\mathrm{RV}} \right )$ of the
two epochs are also given in the same table. The $\overline{\mathrm{RV}}$
values of the objects in UScoOB are consistent with the earlier measurements
of \citet{muzerolle:2003} in Table~\ref{tab:literatures}.

@@

%
% This may be out of place =======================
An alternative explanation for the RV variations found in the binary
candidates above is a type of stellar pulsations found by
\citet{palla:2005}. They studied the non-adiabatic, linear
instability of very VLM stars and BD during the deturium burning phase
in the core, and found unstable fundamental modes in the time-scale
between $\sim 1$ and $\sim 5$~h for the object mass from $0.02$ and
$0.1\,\Msun$.  By using their pulsation periods and by assuming that the
pulsation amplitudes is 5~per~cent of the radius of the objects, we
find that the expected RV variation can be $>6\,\kmps$.
As mentioned in Section~\ref{sec:observation-reduction}, for each object at a
given night, two separate spectra are obtained consecutively (with
time separations of $\sim 0.5$~h.). Before co-adding the two spectra
together, we have measured the RV from each spectrum to check if there
is a large jump in the RV measurements in short
time-scale. The shifts in the RV values in measured in the two
consecutive spectra are found to be between $XXX$ and  $XXX\,\kmps$,
which is much smaller than the expected RV variation based on the
models of \citet{palla:2005}; hence, found no/little indication of the
pulsations.

@@

\begin{center}
\begin{tabular}{lclc}
\hline 
&
$\sigma_{\mathrm{sys}}=0.0\,\kmps$&
&
$\sigma_{\mathrm{sys}}=0.2\,\kmps$\tabularnewline
obj.~ID&
$p$&
obj.~ID&
$p$\tabularnewline
\hline
USco~104&
$4.3\times10^{-12}$&
USco~104&
$3.5\times10^{-6}$\tabularnewline
GY~310&
$1.8\times10^{-10}$&
GY~310&
$2.0\times10^{-5}$\tabularnewline
GY~141&
$1.6\times10^{-4}$&
$\cdots$&
$\cdots$\tabularnewline
USco~55&
$7.2\times10^{-4}$&
GY~141&
$5.8\times10^{-2}$\tabularnewline
$\cdots$&
$\cdots$&
USco~55&
$1.6\times10^{-1}$\tabularnewline
USco~101&
$8.9\times10^{-2}$&
USco~130&
$2.3\times10^{-1}$\tabularnewline
USco~66&
$9.2\times10^{-2}$&
USco~132&
$2.6\times10^{-1}$\tabularnewline
USco~75&
$2.8\times10^{-1}$&
USco~101 &
$2.8\times10^{-1}$\tabularnewline
USco~109&
$3.0\times10^{-1}$&
GY~5&
$2.8\times10^{-1}$\tabularnewline
USco~112&
$3.1\times10^{-1}$&
USco~75&
$3.9\times10^{-1}$\tabularnewline
USco~132&
$3.2\times10^{-1}$&
USco~66&
$4.3\times10^{-1}$\tabularnewline
USco-~30&
$3.2\times10^{-1}$&
USco~128&
$4.6\times10^{-1}$\tabularnewline
GY~5&
$3.7\times10^{-1}$&
USco~53&
$5.8\times10^{-1}$\tabularnewline
USco~53 &
$4.2\times10^{-1}$&
USco40&
$6.0\times10^{-1}$\tabularnewline
USco~67&
$4.8\times10^{-1}$&
USco~112&
$6.5\times10^{-1}$\tabularnewline
USco~128&
$5.4\times10^{-1}$&
USco~100&
$6.6\times10^{-1}$\tabularnewline
USco~100&
$6.2\times10^{-1}$&
USco~67&
$7.0\times10^{-1}$\tabularnewline
USco~40&
$7.2\times10^{-1}$&
USco~109&
$7.1\times10^{-1}$\tabularnewline
\hline
\end{tabular}

>>>>

The radial velocities of each object were determined by using the cross-correlation
function of the object spectrum with that of a template star which has
a similar spectra type. By visual inspection,
the wavelength ranges used for the cross-correlation calculations
are chosen by avoiding the regions of spectra affected by the telluric
lines, and defects and fringes (in near infrared) of the CCDs. The
radial velocities of objects with respect to the template are obtained
by measuring the location of the peak in the cross-correlation
function. The location of the peak is determined by fitting the
section of the cross-correlation function around the peak by a second
order polynomial. LHS~49 (Proxima Cen, M5.5) was
chosen as the template for this purpose. The heliocentric radial velocity of the
template object LHS~49 was obtained by measuring the wavelength shifts
of the prominent photospheric absorption features \ion{K}{i}~$\lambda\lambda$7664.911,~7698.974.
This gives us $\mathrm{RV_{LHS49}=}-22.6\pm0.5\,\kmps$, which is
in good agreement with the earlier measurement of \citet{garcia-sanchez:2001}
who found $\mathrm{RV_{LHS49}=}-21.7\pm1.8\,\kmps$. The heliocentric
RV of each object can be then calculated by adding $\mathrm{RV_{LHS49}}$
with the RV of each object with respect to LHS~49. In the following
measurements of the heliocentric radial velocities, our measurement
($\mathrm{RV_{LHS49}=}-22.6\pm0.5\,\kmps$) will be used for consistency.

Before applying the cross-correlation technique to our main targets,
we have applied the technique to the radial velocity standard HD~140538
for which an high-accuracy RV measurement via the fixed-configuration,
cross-dispersed échelle spectrograph Elodie \citep{baranne:1996}
is available. This was done so to ensure not only the validity of
the cross-correlation technique, but also the validity of the wavelength
calibration. In this test, we found the heliocentric $\mathrm{RV_{HD\,140538}=}18.18\pm0.62\,\kmps$
which is in good agreement with the Elodie radial velocity measurement
of $19.00\pm0.05\,\mathrm{km\, s^{-1}}$ \citep{udry:1999}.

The result of the heliocentric RV measurements (from two epochs for
each object) is summarised in Table~\ref{tab:summary_results} along
with the uncertainties. The table also lists the uncertainties in
the `relative' radial velocities ($\sigma_{\mathrm{RRV}}$) which
is the uncertainty of RV with respect to the template star. 
To determine the $\sigma_{\mathrm{RRV}}$, we have used the following
procedure: (1)~add random gaussian deviate noise to the object spectra using the
corresponding variance spectra, (2)~compute the cross-correlation
curve using the spectra with added noise and the template, (3)~measure
the RV by location the peak, (4)~repeat 1--3 for 100 times, and computes the standard
deviation of the 100 RV values. This should provide a good estimate of
uncertainties in RV cased by the uncertainties in the flux levels.
%The two consecutive measurements of RVs (from same nights) are used to
%find the (absolute/heliocentric) RV of the night, and their uncertainties are obtained from
%the standard deviations of the mean ($\sigma_{\mathrm{RRV}}$) and
%the uncertainty of the template RV value (c.f.~\citealt{joergens:2006}).
Note that for finding a shift in the radial velocities from two
different epochs, only $\sigma_{\mathrm{RRV}}$ is important since one
only requires relative (with respect to a template) RV values.
The `average' radial velocities $\left (\overline{\mathrm{RV}} \right )$ of the
two epochs are also given in the same table. The $\overline{\mathrm{RV}}$
values of the objects in UScoOB are consistent with the earlier measurements
of \citet{muzerolle:2003} in Table~\ref{tab:literatures}.

@@

%
% This may be out of place =======================
An alternative explanation for the RV variations found in the binary
candidates above is a type of stellar pulsations found by
\citet{palla:2005}. They studied the non-adiabatic, linear
instability of very VLM stars and BD during the deturium burning phase
in the core, and found unstable fundamental modes in the time-scale
between $\sim 1$ and $\sim 5$~h for the object mass from $0.02$ and
$0.1\,\Msun$.  By using their pulsation periods and by assuming that the
pulsation amplitudes is 5~per~cent of the radius of the objects, we
find that the expected RV variation can be $>6\,\kmps$.
As mentioned in Section~\ref{sec:observation-reduction}, for each object at a
given night, two separate spectra are obtained consecutively (with
time separations of $\sim 0.5$~h.). Before co-adding the two spectra
together, we have measured the RV from each spectrum to check if there
is a large jump in the RV measurements in short
time-scale. The shifts in the RV values in measured in the two
consecutive spectra are found to be between $0.02\pm0.10$ (USco~128)
and  $2.04\pm1.67\,\kmps$ (USco~66), which is much smaller than the
expected RV variation based on the models of \citet{palla:2005};
hence, found little indication of the pulsations.

@@

\begin{center}
\begin{tabular}{lclc}
\hline 
&
$\sigma_{\mathrm{sys}}=0.0\,\kmps$&
&
$\sigma_{\mathrm{sys}}=0.2\,\kmps$\tabularnewline
obj.~ID&
$p$&
obj.~ID&
$p$\tabularnewline
\hline
USco~104&
$4.3\times10^{-12}$&
USco~104&
$3.5\times10^{-6}$\tabularnewline
GY~310&
$1.8\times10^{-10}$&
GY~310&
$2.0\times10^{-5}$\tabularnewline
GY~141&
$1.6\times10^{-4}$&
$\cdots$&
$\cdots$\tabularnewline
USco~55&
$7.2\times10^{-4}$&
GY~141&
$5.8\times10^{-2}$\tabularnewline
$\cdots$&
$\cdots$&
USco~55&
$1.6\times10^{-1}$\tabularnewline
USco~101&
$8.9\times10^{-2}$&
USco~130&
$2.3\times10^{-1}$\tabularnewline
USco~66&
$9.2\times10^{-2}$&
USco~132&
$2.6\times10^{-1}$\tabularnewline
USco~75&
$2.8\times10^{-1}$&
USco~101 &
$2.8\times10^{-1}$\tabularnewline
USco~109&
$3.0\times10^{-1}$&
GY~5&
$2.8\times10^{-1}$\tabularnewline
USco~112&
$3.1\times10^{-1}$&
USco~75&
$3.9\times10^{-1}$\tabularnewline
USco~132&
$3.2\times10^{-1}$&
USco~66&
$4.3\times10^{-1}$\tabularnewline
USco~130&
$3.2\times10^{-1}$&
USco~128&
$4.6\times10^{-1}$\tabularnewline
GY~5&
$3.7\times10^{-1}$&
USco~53&
$5.8\times10^{-1}$\tabularnewline
USco~53 &
$4.2\times10^{-1}$&
USco~40&
$6.0\times10^{-1}$\tabularnewline
USco~67&
$4.8\times10^{-1}$&
USco~112&
$6.5\times10^{-1}$\tabularnewline
USco~128&
$5.4\times10^{-1}$&
USco~100&
$6.6\times10^{-1}$\tabularnewline
USco~100&
$6.2\times10^{-1}$&
USco~67&
$7.0\times10^{-1}$\tabularnewline
USco~40&
$7.2\times10^{-1}$&
USco~109&
$7.1\times10^{-1}$\tabularnewline
\hline
\end{tabular}
